---
title: "Nonparametric Project of Agricultural Productivity in the U.S."
subtitle: "Permutational Tests "
author:
    - "Sofia Moroni^[sofia.moroni@mail.polimi.it]"
output:
    pdf_document:
        toc: true
        toc_depth: 3
        number_section: true
        #keep_md: TRUE
    html_document:
        toc: true
        toc_float: true
        number_sections: true
date: "2023-06-27"
editor_options:
    chunk_output_type: inline
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    dev = c('pdf'),
    fig.align = 'center',
    fig.path = 'output/',
    fig.height = 3,
    fig.width = 6
)
```

# Load libraries and data

```{r, message=FALSE}
library(pbapply)
library(mgcv)
```

```{r echo = T, results = 'hide'}
data_path = file.path('data')
output_path = file.path('results')
data =
    read.table(
        file.path(data_path, 'agricultural_indices.csv'),
        header = T,
        sep = ';'
    )

# Sostituzione delle virgole con punti
data<- data.frame(lapply(data, function(x) gsub(",", ".", x)))
data <- as.data.frame(lapply(data, as.numeric))
data<- data[1:68,]

set.seed(100)
B = 1000
n = nrow(data)
```

Studying the relationship between input variables and output variable

```{r}
library(car)
with(data  , scatterplotMatrix(data.frame(Total.agricultural.output, Capital.Durable.equipment.Input, Capital.Service.buildings.Input)))

with(data ,scatterplotMatrix(data.frame(Total.agricultural.output,Capital.Land.Input,Capital.Inventories.Input)))

with(data , scatterplotMatrix(data.frame(Total.agricultural.output,Labor.Self.employed.and.unpaid.family.Input,LaborHired.labor.Input)))

with(data, scatterplotMatrix(data.frame(Total.agricultural.output,Intermediate.Feed.and.seed.Input,Intermediate.Energy.Input,Intermediate.Fertilizer.and.lime.Input)))
                                        
with(data, scatterplotMatrix(data.frame(Total.agricultural.output,Intermediate.Pesticides.Input,Intermediate.Purchased.services.Input)))

```

Starting with all the covariates
```{r}
model_gam = gam(
    data$Total.agricultural.output ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + data$Capital.Land.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Feed.and.seed.Input
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Fertilizer.and.lime.Input
    + data$Intermediate.Purchased.services.Input
   )

summary(model_gam)
plot(model_gam)
```


# $H_0$: `Feed and seed` $= 0$ VS $H_1$: `Feed and seed` $\neq 0$

```{r}
T0 = abs(summary(model_gam)$p.table[5, 3])
gam.H0 = gam(
    data$Total.agricultural.output ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + data$Capital.Land.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Fertilizer.and.lime.Input
    + data$Intermediate.Purchased.services.Input
   )

    
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + data$Capital.Land.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Feed.and.seed.Input
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Fertilizer.and.lime.Input
    + data$Intermediate.Purchased.services.Input
   )

    return(abs(summary(gam.perm)$p.table[5, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```
```{r}

hist(sort(T_H0)[-1000],
     breaks = 100,
     col = 'white',
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)



plot(ecdf(sort(T_H0)[-1000]), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)


P = sum(T_H0 >= T0) / B
P
```
Can accept H0, so I remove the variable


# $H_0$: `Fertilizer and lime` $= 0$ VS $H_1$: `Fertilizer and lime` $\neq 0$

```{r}
model_gam = gam(data$Total.agricultural.output ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + data$Capital.Land.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Fertilizer.and.lime.Input
    + data$Intermediate.Purchased.services.Input
   )

summary(model_gam)

T0 = abs(summary(model_gam)$p.table[7, 3])
gam.H0 = gam(data$Total.agricultural.output ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + data$Capital.Land.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Purchased.services.Input
   )
    
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + data$Capital.Land.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Fertilizer.and.lime.Input
    + data$Intermediate.Purchased.services.Input
    )
    return(abs(summary(gam.perm)$p.table[7, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```
```{r}
hist(sort(T_H0)[-1000],
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(sort(T_H0)[-1000]), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```
Can accept H0, so I remove the variable `Feed and Seed`


# $H_0$: `Land` $= 0$ VS $H_1$: `Land` $\neq 0$

```{r}
model_gam = gam(data$Total.agricultural.output ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + data$Capital.Land.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Purchased.services.Input
   )

summary(model_gam)

T0 = abs(summary(model_gam)$p.table[3, 3])
gam.H0 = gam(data$Total.agricultural.output ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Purchased.services.Input
   )
    
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + data$Capital.Land.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Purchased.services.Input
   )
    return(abs(summary(gam.perm)$p.table[3, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```
```{r}

hist(sort(T_H0)[-1000],
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     col = 'white',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)


plot(ecdf(sort(T_H0)[-1000]), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```

Can accept H0, so I remove the variable 


# $H_0$: `Purchased Services` $= 0$ VS $H_1$: `Purchased Services` $\neq 0$

```{r}
model_gam = gam(data$Total.agricultural.output ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Purchased.services.Input
   )

summary(model_gam)

T0 = abs(summary(model_gam)$p.table[6, 3])
gam.H0 = gam(data$Total.agricultural.output ~ s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Purchased.services.Input
   )
    
res.H0 = gam.H0$residuals

wrapper = function() {
    permutation = sample(n)
    res.H0.perm = res.H0[permutation]
    Y.perm.H0 = gam.H0$fitted + res.H0.perm
    gam.perm = gam(
        Y.perm.H0 ~  s(data$Capital.Durable.equipment.Input, bs = 'cr')
    + data$Capital.Service.buildings.Input
    + s(data$Capital.Inventories.Input,bs = 'cr')
    + data$Labor.Self.employed.and.unpaid.family.Input
    + s(data$LaborHired.labor.Input, bs = 'cr')
    + data$Intermediate.Energy.Input
    + data$Intermediate.Pesticides.Input
    + data$Intermediate.Purchased.services.Input
   )
    return(abs(summary(gam.perm)$p.table[6, 3]))
}
T_H0 = pbreplicate(B, wrapper(), simplify = 'vector')
```
```{r}
hist(sort(T_H0)[-1000],
     breaks = 100,
     main = 'Permutational distribution of test statistics',
     xlab = '')
abline(v = T0, col = 'red', lwd = 4)

plot(ecdf(sort(T_H0)[-1000]), main = 'ECDF of test statistics')
abline(v = T0, col = 'red', lwd = 4)
P = sum(T_H0 >= T0) / B
P
```








